---
title: "Introduction"
subtitle: "Summer School JGU Mainz — Advanced Methods in Behavioral Economics, 2021"
author: "Carina I. Hausladen"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
rm(list=ls())
#knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(viridis)
library(hrbrthemes)
library(readr)
```

```{r read, include=FALSE}
df <- read_csv("SS_skills.csv",skip = 1)
df2<-df[,18:35]
df2<-df2[-1,-c(10,12)]
```

## Programming Language
```{r language, echo=FALSE, fig.align='center'}
df2 %>% count(`python (3)`,`R/RStudio`) %>% 
  drop_na() %>%
  pivot_longer(cols = `python (3)`:`R/RStudio`,names_to = c("language"),values_to = "choice")%>%
  ggplot(., aes(fill=choice, y=n, x=language)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T) +
    ggtitle("Python and R") +
    theme_ipsum() +
    xlab("")
```

## Where to code?
```{r prefer, echo=FALSE, fig.align='center'}
df2 %>% count(`I prefer...`)%>%
  drop_na() %>%
    ggplot(., aes(x=`I prefer...`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("I prefer") +
    theme_ipsum() +
    xlab("")
``` 

## Where to find code?
```{r git, echo=FALSE, fig.align='center'}
df2 %>% count(`Git (version control)`)%>%
  drop_na() %>%
    ggplot(., aes(x=`Git (version control)`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("Git") +
    theme_ipsum() +
    xlab("")
```

## How to install?
```{r venv_pip, echo=FALSE, fig.align='center'}
df2 %>% count(`I use... - venv`,`I use... - pip`)%>% 
  drop_na() %>%
  pivot_longer(cols = `I use... - venv`:`I use... - pip`, names_to = c("language"),values_to = "choice")%>%
  ggplot(., aes(fill=choice, y=n, x=language)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T) +
    ggtitle("Venv/pip") +
    theme_ipsum() +
    xlab("")
``` 


# Session 1

## Political Speech

`r emo::ji("chart_with_upwards_trend")` [Plenarprotokolle Deutscher Bundestag](https://www.bundestag.de/dokumente/protokolle/plenarprotokolle), [New York Times front pages](http://www.amber-boydstun.com/supplementary-information-for-making-the-news.html), 
[Manifestos Project Database](https://manifestoproject.wzb.eu)

* Programming language: R (mainly [quanteda](https://quanteda.io))
* Methods
    * Descriptive Methods `r emo::ji("teacher")` 12', `r emo::ji("technologist")` 12'
      * Corpus, types, tokens, sentences
      * Readability scores ([overview](https://quanteda.io/reference/textstat_readability.html), [Flesch (1948)](https://psycnet.apa.org/record/1949-01274-001))
      * Dfm
      * [Comparing texts](https://quanteda.io/reference/textstat_simil.html) (e.g. via cosine similarity)
      * Keywords in context
    * Classification `r emo::ji("teacher")` 12', `r emo::ji("technologist")` 12'
      * [Train/Test split](https://towardsdatascience.com/train-validation-and-test-sets-72cb40cba9e7) 
      * Naive Bayes Classifier ([Schütze et al. (2008), ch 13](https://nlp.stanford.edu/IR-book/pdf/irbookonlinereading.pdf))
      * Importance of sampling
      * Evaluation metrics
      
---

  * Scaling models `r emo::ji("teacher")` 12', `r emo::ji("technologist")` 12'
    * Wordscore ([Laver et al. (2003)](https://www.cambridge.org/core/journals/american-political-science-review/article/extracting-policy-positions-from-political-texts-using-words-as-data/4F4676E80A79E01EAAB88EF3F2D1B733)) 
    * Wordfish ([Slapin (2008)](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1540-5907.2008.00338.x))
    * Wordshoal ([Lauderdale (2016)](https://www.cambridge.org/core/journals/political-analysis/article/measuring-political-positions-from-legislative-speech/35D8B53C4B7367185325C25BBE5F42B4))
  * Measuring divisiveness in political speech ([Ash (2017)](https://www.journals.uchicago.edu/doi/full/10.1086/692587?af=R&mobileUi=0)) `r emo::ji("teacher")` 12', `r emo::ji("technologist")` 12'

 
```{r some_word, echo=FALSE, fig.align='center'}
df2 %>% count(`Some topics - wordscore/wordfish/wordshoal`)%>%
  drop_na() %>%
  ggplot(., aes(x=`Some topics - wordscore/wordfish/wordshoal`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("Scaling Models") +
    theme_ipsum() +
    xlab("")
``` 

    
#  Session 2 

## Tax Risk Management (I)
`r emo::ji("chart_with_upwards_trend")` Tax reports from German companies listed in the STOXX Europe 600.

* Programming language: Python (mostly [scikit-learn](https://scikit-learn.org/stable/))
* Methods
  * `r emo::ji("teacher")` 10', `r emo::ji("technologist")` 10'
    * Text preprocessing ([Stemming, Lemmatization](https://nlp.stanford.edu/IR-book/html/htmledition/stemming-and-lemmatization-1.html))
    * Train/Test split
    * [Feature extraction](https://scikit-learn.org/stable/modules/feature_extraction.html)
    
---

  * Classification `r emo::ji("teacher")` 10', `r emo::ji("technologist")` 10'
    * [Multinomial Bayes](https://scikit-learn.org/stable/modules/naive_bayes.html#multinomial-naive-bayes) 
    * [Passive Aggressive Algorithms](https://scikit-learn.org/stable/modules/linear_model.html#passive-aggressive-algorithms)
    * [Support Vector Machines](https://scikit-learn.org/stable/modules/svm.html#classification)
    * Gazetteer/Entity Dictionary

```{r some_class, echo=FALSE, fig.align='center'}
df2 %>% count(`Some topics - classification, e.g. KNN, SVM, RF, NB, NN, XGBoost`)%>%
  drop_na() %>%
  ggplot(., aes(x=`Some topics - classification, e.g. KNN, SVM, RF, NB, NN, XGBoost`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("Classification, e.g. KNN, SVM, RF, NB, NN") +
    theme_ipsum() +
    xlab("")
```  

## Predicting (Non) Compliance

`r emo::ji("chart_with_upwards_trend")` chat data from a [laboratory experiment](https://www.socialscienceregistry.org/trials/5049)

* Programming language: Python (mostly [scikit-learn](https://scikit-learn.org/stable/))
* Presentation: Chapter 2 from my [dissertation](https://kups.ub.uni-koeln.de/12231/) `r emo::ji("teacher")` 15'

# Session 3
## Predicting (Non) Compliance—Continued

* Methods
  * Feature engineering `r emo::ji("teacher")` 10', `r emo::ji("technologist")` 10'
    * Input: concatenating chat on different levels (individuals, versus groups)
    * Output: applying different thresholds to binarize the continuous dependent variable
    * Sampling ([oversampling](https://imbalanced-learn.org/stable/over_sampling.html#random-over-sampler))
    * [Stratified Group Shuffle Split](https://stackoverflow.com/questions/56872664/complex-dataset-split-stratifiedgroupshufflesplit)
  * Embeddings `r emo::ji("teacher")` 15', `r emo::ji("technologist")` 15'
    * Word2Vec ([Mikolov (2013)](https://arxiv.org/abs/1301.3781))
    * GloVe ([Pennington (2014)](https://aclanthology.org/D14-1162.pdf))
    * fastText ([Joulin (2016)](https://arxiv.org/pdf/1607.01759.pdf?fbclid=IwAR1wttEXho2gqk3BasKDuncgftN5I5lmH2TbIgvGuHxfutM3IavbateHH9A))
    * Doc2Vec (PV-DM) ([Le (2014)](http://proceedings.mlr.press/v32/le14.html))

---

  * Classification `r emo::ji("teacher")` 15', `r emo::ji("technologist")` 15'
    * [Tuning Hyperparametes](https://scikit-learn.org/stable/modules/grid_search.html#grid-search)
    * [Cross-validation](https://scikit-learn.org/stable/modules/cross_validation.html#cross-validation)
    * Models
      * [Linear classifier (log loss) with SGD training](https://scikit-learn.org/stable/modules/sgd.html#sgd)
      * [KNN](https://scikit-learn.org/stable/modules/neighbors.html#classification)
 
---

  * [ensemble methods](https://scikit-learn.org/stable/modules/ensemble.html)
    * [Random Forest](https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.RandomForestClassifier.html)
    * [bagging classifier](https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.BaggingClassifier.html)
    * [stacked generalization](https://scikit-learn.org/stable/modules/ensemble.html#stacked-generalization)
        
```{r some_bagging, echo=FALSE, fig.align='center'}
df2 %>% count(`Some topics - bagging/boosting`)%>%
  drop_na() %>%
  ggplot(., aes(x=`Some topics - bagging/boosting`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("") +
    theme_ipsum() +
    xlab("")
``` 

# Session 4
## Predicting (Non) Compliance—Continued

* Parallelization `r emo::ji("technologist")` 5'
* Scoring `r emo::ji("teacher")` 10', `r emo::ji("technologist")` 10'
  * [Accuracy, precision, recall, f1-score, AUC](https://scikit-learn.org/stable/modules/model_evaluation.html#precision-recall-f-measure-metrics)
  * Evaluate within and across datasets ([Straube (2014)](https://www.frontiersin.org/articles/10.3389/fncom.2014.00043/full))
    
```{r some_parallel, echo=FALSE, fig.align='center'}
df2 %>% count(`Some topics - parallel processing`)%>%
  drop_na() %>%
  ggplot(., aes(x=`Some topics - parallel processing`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("Parallel processing") +
    theme_ipsum() +
    xlab("")
``` 

## Ideological Direction in Judicial Opinion
`r emo::ji("chart_with_upwards_trend")` US Circuit Court judicial decisions

* Programming language: Python (mostly [scikit-learn](https://scikit-learn.org/stable/))
* Presentation: [Hausladen et al. 2020](https://www.sciencedirect.com/science/article/pii/S0144818819303667) `r emo::ji("teacher")` 15'
* Methods
  * Feature engineering: exploiting the specific structure of judicial texts (citations/quotations)
  * [Probability calibration](https://scikit-learn.org/stable/modules/calibration.html#) `r emo::ji("teacher")` 5', `r emo::ji("technologist")` 10'
  
```{r some_probcal, echo=FALSE, fig.align='center', fig.height=3}
df2 %>% count(`Some topics - probability calibration`)%>%
  drop_na() %>%
  ggplot(., aes(x=`Some topics - probability calibration`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("Probability calibration") +
    theme_ipsum() +
    xlab("")
``` 

## Further Techniques
* Sentiment Analysis with LSTM `r emo::ji("teacher")` 7', `r emo::ji("technologist")` 7'
* RCNN ([Lai et al. (2015)](https://www.aaai.org/ocs/index.php/AAAI/AAAI15/paper/viewPaper/9745)) `r emo::ji("teacher")` 7', `r emo::ji("technologist")` 7'

```{r some_senti, echo=FALSE, fig.align='center'}
df2 %>% count(`Some topics - sentiment analysis`)%>%
  drop_na() %>%
  ggplot(., aes(x=`Some topics - sentiment analysis`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("Sentiment Analysis") +
    theme_ipsum() +
    xlab("")
``` 

# Session 5

## Charting the Type Space

`r emo::ji("chart_with_upwards_trend")` simulated decision data from a multi-stage public good game

* Programming language: R (mostly [dtw clust](https://cran.r-project.org/web/packages/dtwclust/dtwclust.pdf))
* Presentation: Chapter 3 of my [dissertation](https://kups.ub.uni-koeln.de/12231/) `r emo::ji("teacher")` 15'
* Methods ([Sarda-Espinosa (2017)](https://mran.microsoft.com/snapshot/2018-07-24/web/packages/dtwclust/vignettes/dtwclust.pdf)) `r emo::ji("teacher")` 20', `r emo::ji("technologist")` 20'
    * distance measures (dtw, sdtw)
    * prototypes (dba)
    * clustering algorithms (hierarchical, partitional)
    * cluster evaluation (e.g. majority vote of different indices)

```{r some_clust, echo=FALSE, fig.align='center', fig.height=3}
df2 %>% count(`Some topics - clustering, e.g. cluster evaluation/centroid functions/distance measures`)%>%
  drop_na() %>%
  ggplot(., aes(x=`Some topics - clustering, e.g. cluster evaluation/centroid functions/distance measures`, y=n, fill=factor(n)))+
              geom_bar(stat='identity')+
    scale_fill_viridis(discrete = T) +
    ggtitle("") +
    theme_ipsum() +
    xlab("")
``` 
    
## Tax Risk Management (II)

`r emo::ji("chart_with_upwards_trend")` Tax reports from German companies listed in the STOXX Europe 600.

* Programming language: Python (mostly [scikit-learn](https://scikit-learn.org/stable/))
* Methods `r emo::ji("teacher")` 10', `r emo::ji("technologist")` 10'
  * Clustering ([K-means](https://scikit-learn.org/stable/modules/clustering.html#k-means))
  * Visualize high-dimensional data ([t-SNE](https://scikit-learn.org/stable/modules/manifold.html#t-sne))
  * Topic Models ([NFM](https://scikit-learn.org/stable/modules/decomposition.html#non-negative-matrix-factorization-nmf-or-nnmf), [LDA](https://scikit-learn.org/stable/modules/decomposition.html#latent-dirichlet-allocation-lda))

# Let's compete for Swiss chocolate! 
After every session, it's quiz time.
`r emo::ji("fox")` 
  