---
title: "Classification"
subtitle: "Summer School JGU Mainz — Advanced Methods in Behavioral Economics, 2021"
author: "Carina I. Hausladen"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
bibliography: ss21.bib 
---

```{r setup, include=FALSE, cache=TRUE}
rm(list = ls())
library(quanteda)
library(quanteda.textmodels)
library(caret)
library(devtools)
library(quantedaData)
library(ggplot2)
library(readtext)
library(XML)
library(stringr)
library(lubridate)
library(sampling)
library(reshape2)
library(knitr)
library(e1071)
```

## Methods
In this tutorial you will learn about:

* Train/Test split
* Naive Bayes Classifier 
* Importance of sampling
* Evaluation metrics

## Data
- We use a corpus of article titles and descriptions from the [New York Times front pages](http://www.amber-boydstun.com/supplementary-information-for-making-the-news.html). 
- Our classification goal is to assign the story to a class of subject category.

```{r df, include=FALSE, cache=TRUE}
nytfrontpage <- readtext("data/nytfrontpage.csv")
summary(nytfrontpage, 5)
```

```{r construct_corpus, include=FALSE, cache=TRUE}
#The category "Arts and Entertainment" (category 23) only has 4 entries and is therefore excluded from the corpus. If I included this category, the classifier would assign high weights to those words that are in the texts assigned to this small category. As soon as the same words appear in a text that is assigned to another category the classifier would be very likely to assign this text to the wrong category. 

topicareas <- as.numeric(names(table(nytfrontpage$majortopic)[table(nytfrontpage$majortopic)>10]))
nytfrontpage2 <- subset(nytfrontpage, nytfrontpage$majortopic %in% topicareas)
nytfrontpage2$majortopic <- as.factor(nytfrontpage2$majortopic)

text<-paste(nytfrontpage2$title, nytfrontpage2$description, sep=" ")
names<-c("Macroeconomics", "Civil Rights","Health","Agriculture","Labor",
         "Education","Environment", "Energy","Immigration", "Transportation",
         "Law","Social Welfare","Community","Banking","Defense", "Space",
         "Trade","International Affairs","Government Operations",
         "Public Lands","Administration","Weather","Fires","Undefined","Sports",
         "Death","Churches", "Other")
nytfrontpage2$majortopic.text<-factor(nytfrontpage2$majortopic, 
                                     levels = sort(unique(nytfrontpage2$majortopic)), 
                                     labels = names)

corpus.nyt<-corpus(text)
docvars(corpus.nyt,"doc_id")          <- 1:dim(nytfrontpage2)[1]
docvars(corpus.nyt,"majortopic")      <- nytfrontpage2$majortopic
docvars(corpus.nyt,"majortopic.text") <- nytfrontpage2$majortopic.text
```

```{r dscrb, echo=FALSE, out.width='.32\\linewidth', fig.width=5, fig.height=1.85, fig.show='hold',fig.align='center', cache=TRUE}
y<-table(docvars(corpus.nyt,"majortopic.text"))
ggplot(as.data.frame(y), 
       aes(x = reorder(Var1, Freq), y = Freq)) + geom_bar(stat="identity")+ theme(axis.text.x = element_text(angle = 90, hjust = 1, size=6)) + theme(axis.text.y = element_text(size=6)) + theme(axis.title=element_text(size=6))  + labs(x = "Categories", y = "Absolute Frequency", size=6)
```

The corpus consists of `r dim(nytfrontpage)[1]` documents and `r length(unique(nytfrontpage$majortopic))` topics.

## Train-test split

[Jason Prownlee](https://machinelearningmastery.com/train-test-split-for-evaluating-machine-learning-algorithms/)

- The train-test split is a technique for evaluating the performance of a machine learning algorithm.
- It can be used for classification or regression problems and can be used for any supervised learning algorithm.
- Train Dataset: Used to fit the machine learning model.
- Test Dataset: Used to evaluate the fit machine learning model. The objective is to estimate the performance of the machine learning model on new data: data not used to train the model.

In our example: The trainingset consists of 3584 randomly sampled documents.
```{r train, include=TRUE, cache=TRUE}
set.seed(42)
train <- sample(1:5000, 3584, replace = FALSE)
head(train, 10)

3584/dim(nytfrontpage)[1]
```

```{r train_tokens, include=FALSE, error=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
# create docvar with ID
corpus.nyt$id_numeric <- 1:ndoc(corpus.nyt)

# tokenize texts
toks_nyt <- tokens(corpus.nyt, remove_punct = TRUE, remove_number = TRUE) %>% 
               tokens_remove(pattern = stopwords("en")) %>% 
               tokens_wordstem()
dfmt_nyt <- dfm(toks_nyt)

# get training set
dfmat_training <- dfm_subset(dfmt_nyt, id_numeric %in% train)

# get test set (documents not in id_train)
dfmat_test <- dfm_subset(dfmt_nyt, !id_numeric %in% train)
```

## Naive Bayes Classifier

### Theory
@Jurafsky2018

- A Naive Bayesian classifier  makes a simplifying (naive) assumption about how the features interact.
- We represent a text document as if it was a bag-of-words.

---

### Train the model
[quanteda tutorial](https://tutorials.quanteda.io/machine-learning/nb/)

Next, we train the Naive Bayes classifier.
```{r nb, include=TRUE, error=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
tmod_nb <- textmodel_nb(dfmat_training, dfmat_training$majortopic)
```


## Evaluation

### Confusion Matrix

- The diagonal elements of a confusion matrix represent the number of points for which the predicted label is equal to the true label.
- The off-diagonal elements are those that are mislabeled by the classifier. 
- The higher the diagonal values of the confusion matrix the better.

```{r cnfsn_mtrx, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
dfmat_matched <- dfm_match(dfmat_test, features = featnames(dfmat_training))
actual_class <- dfmat_matched$majortopic
predicted_class <- predict(tmod_nb, newdata = dfmat_matched)
conftable <- table(actual_class, predicted_class)
conftable[1:10,1:10]
```

```{r cnfsn_mtrx_kable, include=FALSE, error=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
#We can also use the function `confusionMatrix()` to assess the performance of the classification.
kable(round(confusionMatrix(conftable)[["overall"]], digits=2))
```

## Metrics

- In order to evaluate the classifier, we calculate the metrics accuracy, precision and recall.
- Accuracy measures the fraction of all correct predictions.
- Precision: Out of the documents that the model predicts as category "Defense", what is the fraction actually in the category "Defense"? 
- Recall: What is fraction of documents in the category "Defense" that is correctly identified? 

```{r calc_mtrcs, include=FALSE, cache=TRUE}
Accuracy = sum(diag(conftable))/sum(conftable)

Precision<-c()
for(i in 1:dim(conftable)[1]){
  Precision[i] = diag(conftable)[i]/sum(conftable[,i])
}

Recall<-c()
for(i in 1:dim(conftable)[1]){
  Recall[i] = diag(conftable)[i]/sum(conftable[i,], na.rm = TRUE)
}

Macro_average_precision = mean(Precision, na.rm = TRUE)  
Macro_average_recall = mean(Recall, na.rm = TRUE)  
```

```{r names2, include=FALSE, cache=TRUE}
names2<-c("Macroeconomics","Civil Rights","Health","Agriculture","Labor and Employment","Education","Environment","Energy","Immigration","Transportation","Law, Crime","Social Welfare","Community Development","Banking","Defense","Space, Science, Technology","Foreign Trade","International Affairs","Government Operations","Public Lands","State and Local Government","Weather","Fires","Other","Sports and Recreation","Death Notices","Churches and Religion","Other, Miscellaneous")
```

```{r print_mtrcs, include=FALSE, warning=FALSE, cache=TRUE}
nv<-as.data.frame(cbind(c("Accuracy", "Macro Average Perecision", "Macro Average Recall"),
                          round(c(Accuracy, Macro_average_precision, Macro_average_recall), digits = 2)))
colnames(nv)<-c("Metric", "Value")
kable(nv)
```

```{r plt_mtrcs, echo=FALSE, fig.align='center', fig.height=3.5, cache=TRUE}
d<-as.data.frame(cbind(as.numeric(rownames(conftable)), as.numeric(Precision), as.numeric(Recall)))
colnames(d)<-c("Class","Precision", "Recall")
d$majortopic.text<-factor(d$Class, levels = sort(d$Class), labels = names2)
d.long<-melt(d, id.vars=c("Class", "majortopic.text"))

ggplot(d.long, aes(fill=variable, y=value, x=reorder(majortopic.text, value))) + 
  geom_bar(position="dodge", stat="identity")+ 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  xlab("")
  
```


# Investigate poor classification

```{r poor_clf, include=FALSE, cache=TRUE}
## The table below shows those categories that perform below both: macro average precision and recall. 
d[,c(2,3)]<-round(d[,c(2,3)], digits=2)
d2<-d[d$Recall<Macro_average_recall & d$Precision<Macro_average_precision,c(2:4)]
colnames(d2)[3]<-"Topic"
kable(d2[order(d2$Precision),],row.names=FALSE)
```

## Correlation with sample size
In order to find out why the classifier performs poorly for certain classes, we plot the share of category in the whole sample size against precision/recall. 
```{r correlation, echo=FALSE,fig.align='center', cache=TRUE}
ss<-as.data.frame(as.matrix(table(docvars(dfmat_training, "majortopic"))))
d$samplesize <- ss[match(d$Class, rownames(ss)),1]
d$samplesize.share<-c()
for(i in 1:length(d$samplesize)){
  d$samplesize.share[i] <- d$samplesize[i]/ sum(d$samplesize)
}

  {plot(d$samplesize.share, d$Precision, ylab = "Precision", xlab = "Share of Category within Sample", pch=20, main="Precision")
  abline(lm(d$Precision~d$samplesize.share))}
```

```{r correlation_recall, include=FALSE,fig.align='center', cache=TRUE}
  {plot(d$samplesize.share, d$Recall, ylab = "Recall", xlab = "Share of Category within Sample", pch=20, main="Recall")
  abline(lm(d$Recall~d$samplesize.share))}
```

## Is the trainingset balanced?

The barplot shows that the trainingset is heavily unbalanced.
The categories are far away from being uniformly distributed nor are the category sizes in the sample of the same size as in the original data set.

```{r split, echo=FALSE,fig.align='center', cache=TRUE}
tss<-as.data.frame(as.matrix(table(nytfrontpage2$majortopic)/sum(table(nytfrontpage2$majortopic))))
d$shares <- tss[match(d$Class, rownames(tss)),1]
colnames(d)[c(6,7)]<-c("Share Sample", "Share Original")
d.long<-melt(d[,c(4,6,7)], id.vars=c("majortopic.text"))

ggplot(d.long, aes(fill=variable, y=value, x=reorder(majortopic.text, value))) + 
      geom_bar(position="dodge", stat="identity") + 
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("") + ylab("Share")
```


## Generate a balanced subset
We generate two different balanced subsets. 

### Stratification

- Stratification is the process of dividing members of the population into homogeneous subgroups before sampling. 
- Then simple random sampling or systematic sampling is applied within each stratum. 
```{r stratification, include=FALSE, cache=TRUE}
f<-as.data.frame(unique(nytfrontpage2$majortopic))
f$count<-c(1:28)
for(i in 1:dim(f)[1]){ # count how often a specific topic is present in the df
  fd<-f[i,1]
  f[i,2]<-table(nytfrontpage2$majortopic==fd)[2]
}

ss<-min(table(nytfrontpage2$majortopic))*length(table(nytfrontpage2$majortopic)) #sample size: 3584
bins_strat<-as.vector((f$count/sum(f$count))*ss)
```

### Uniform distribution
Second, we assure balance via a uniform distribution.
```{r uniform, include=FALSE, cache=TRUE}
bins_uniform<-rep(min(table(nytfrontpage2$majortopic)), dim(f)[1])
```

```{r both, include=FALSE, cache=TRUE}
bins<-as.data.frame(cbind(bins_strat, bins_uniform))
colnames(bins)<-c("Stratification", "Uniform")
```

## Run NB Classifier with different samples and evaluate

`r emo::ji("nerd")`: Investigate the resulting table and plots and compare them to the outcomes of the randomly sampled training set!

```{r cmpr_sampling, results='asis', include=FALSE, warning=FALSE, message=FALSE, error=FALSE, out.width='.49\\linewidth', fig.width=3, fig.height=3,fig.show='hold',fig.align='center', cache=TRUE}
for(q in 1:2){
  set.seed(42)
  s2<-strata(nytfrontpage2,"majortopic", size=bins[,q], method="srswor")

  #train test split
  corpus.nyt$id_numeric <- 1:ndoc(corpus.nyt)
  toks_nyt <- tokens(corpus.nyt, remove_punct = TRUE, remove_number = TRUE) %>% 
                 tokens_remove(pattern = stopwords("en")) %>% 
                 tokens_wordstem()
  dfmt_nyt <- dfm(toks_nyt)
  dfmat_training <- dfm_subset(dfmt_nyt, id_numeric %in% s2$ID_unit)
  dfmat_test <- dfm_subset(dfmt_nyt, !id_numeric %in% s2$ID_unit)
  
  tmod_nb <- textmodel_nb(dfmat_training, dfmat_training$majortopic)
  dfmat_matched <- dfm_match(dfmat_test, features = featnames(dfmat_training))
  
  #confusion matrix
  actual_class <- dfmat_matched$majortopic
  predicted_class <- predict(tmod_nb, newdata = dfmat_matched)
  conftable <- table(actual_class, predicted_class)
  kable(round(confusionMatrix(conftable)[["overall"]], digits=2))

  #metrics
  Accuracy = sum(diag(conftable))/sum(conftable)
  Precision<-c()
  for(i in 1:dim(conftable)[1]){
    Precision[i] = diag(conftable)[i]/sum(conftable[,i])
  }
  Macro_average_precision = mean(Precision, na.rm = TRUE)
  Recall<-c()
  for(i in 1:dim(conftable)[1]){
    Recall[i] = diag(conftable)[i]/sum(conftable[i,])
  }
  Macro_average_recall = mean(Recall, na.rm = TRUE)

#PLOTS

#Single Values
  nv<-as.data.frame(cbind(c("Accuracy", "Macro Average Perecision", "Macro Average Recall"),
                          round(c(Accuracy, Macro_average_precision, Macro_average_recall), digits=2)))
  colnames(nv)<-c("Metrics", "Value")
  print(kable(nv))

  #plot Precision and Recall
  d<-as.data.frame(cbind(as.numeric(rownames(conftable)), as.numeric(Precision), as.numeric(Recall)))
  colnames(d)<-c("Class","Precision", "Recall")
  d$majortopic.text<-factor(d$Class, levels = sort(d$Class), labels = names2)
  d.long<-melt(d, id.vars=c("Class", "majortopic.text"))
  
  print(ggplot(d.long, aes(fill=variable, y=value, x=reorder(majortopic.text, value))) + 
    geom_bar(position="dodge", stat="identity")+ 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    xlab("")+ 
    geom_hline(yintercept = Macro_average_precision, color="green")+ 
    geom_hline(yintercept = Macro_average_recall, color="blue"))
}
```

## References

# It's your turn! `r emo::ji("technologist")` ~10' 