---
title: "Political Speech and Divisiveness"
subtitle: "Summer School JGU Mainz — Advanced Methods in Behavioral Economics, 2021"
author: "Carina I. Hausladen"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
bibliography: ss21.bib 
---

```{r setup, include=FALSE, cache=TRUE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)

library(quanteda)
library(ggplot2)
library(readtext)
library(readr)
library(dplyr)
library(lfe)
library(stargazer)
library(Rmisc)
library(knitr)
library(vroom)

substrRight <- function(x, n){ #function to extract part of strings
substr(x, nchar(x)-n+1, nchar(x))
}
```


# Political Speech and Divisiveness
"Most citizens want a secure country, safe neighborhoods and good schools. 
These issues do get discussed, of course, but a disproportionate amount of attention goes to issues like abortion, medical marijuana, and other narrow issues that simply do not motivate the great majority." @Fiorina2005, p. 202).


## Introduction 

- @Ash2017 use floor speeches by Members of the U.S. Congress and find that U.S. Senators spend more time on divisive issues when they are up for election. 
- In this module, we will apply the model to the context of German politics. 
- We use speech data from the German Bundestag, whereby we will investigate the impact of the election on September 22nd 2013.

## Research Question and Hypothesis

### Research question: 
Do German politicians use more divisive language as soon as elections are close?

### Hypothesis
Greater posturing through divisive language is expected when electoral concerns are stronger.

## What is divisiveness?
- When voters are uncertain about an incumbent's preferences, there is a pervasive incentive to "posture" by spending too much time on divisive issues (which are more informative about a politician's preferences)
- This is at the expense of time spent on common-values issues (which provide greater benefit to voters). 

## Dataset

- The raw text of the record from the German Bundestag can be obtained from the [official homepage](https://www.bundestag.de/dokumente/protokolle/plenarprotokolle). 
- The text on each page of each document is split into two columns which is not correctly detected by the r function "readtext()". 
[Data in txt-format](https://www.bundestag.de/dokumente/protokolle/plenarprotokolle/plenarprotokolle) is available on the official homepage of the Bundestag only from 2013 to 2018. 
- The website [offenesparlament.de](https://github.com/Datenschule/plpr-scraper/tree/master/data/txt) offers txt-data from 2009 to 2015. 
- As this was the most complete dataset available, I chose to analyze this set of texts.
Parsed and preprocessed .csv files are available [here](https://github.com/bundestag/plpr-scraper) and [here](https://github.com/sebas-seck/plpr-scraper).

```{r prpr_df, echo=FALSE, cache=TRUE}
temp = list.files(path="data/prtkll/18WP/csv",
                  pattern="*.csv")
txt_18 = do.call(rbind, lapply(temp, function(x) read.csv(paste("data/prtkll/18WP/csv/",x,sep=""),
                                                       comment.char="#",
                                                       stringsAsFactors = FALSE)))

temp = list.files(path="data/prtkll/17WP/csv",
                  pattern="*.csv")
txt_17 = do.call(rbind, lapply(temp, function(x) read.csv(paste("data/prtkll/17WP/csv/",x,sep=""),
                                                       comment.char="#",
                                                       stringsAsFactors = FALSE)))
print(dim(txt_17))
print(head(txt_17))
```

## Choosing two parties

- The divisiveness measure used by @Ash2017 is tailored to a two-party system. 
- Unlike the U.S. political system, the German parliament consists of more than two parties, therefore a multidimensional approach to measure divisiveness would be needed. 
- To keep it simple, we'll work with a two-dimensional approach in this tutorial. Therefore, we subset the dataset, including those two parties for which the most speech data is recorded: CDU/CSU and the SPD.
```{r select_speech, include=FALSE, cache=TRUE}
txt2<-txt_17[txt_17$type=="speech",] #select type
```

```{r select_party, include=TRUE, cache=TRUE}
dim(txt2)[1]
txt2<-txt2[txt2$speaker_party =="spd" | txt2$speaker_party =="cducsu",]
dim(txt2)[1]
```

```{r preprocess_df, include=FALSE, cache=TRUE}
txt2<-txt2[,c("text", "speaker_party", "speaker_fp", "wahlperiode", "sitzung")]

txt3<-txt_18[txt_18$type=="speech",] #select type
txt3<-txt3[txt3$speaker_party =="spd" | txt3$speaker_party =="cducsu",] #select party

txt3<-txt3[,c("text", "speaker_party", "speaker_fp", "wahlperiode", "sitzung")] 
txt2<-rbind(txt2,txt3)

txt2<-txt2[complete.cases(txt2), ] #no NAs
```

## The Election Treatment
$E_{it}$ represents the election treatment, which equals $1$ for the first and $4$ for the last year of the "Legislaturperiode" (currently up for election).

election: 27.09.2009 <br>

|$E_{it}=1$ | 10.2009-10.2010 |
|$E_{it}=2$ | 11.2010-10.2011 |
|$E_{it}=3$ | 11.2011-10.2012 |
|$E_{it}=4$ | 11.2012-09.2013 |

election: 22.09.2013 <br>

|$E_{it}=1$ | 10.2013-10.2014 | 
|$E_{it}=2$ | 11.2014-10.2015 | 
|$E_{it}=3$ | 11.2015-10.2016 |  
|$E_{it}=4$ | 11.2016-09.2017 |  

election: 24.09.2017   

```{r year, include=FALSE, cache=TRUE}
#year
  txt2$wahlperiode<-as.numeric(txt2$wahlperiode)
  txt2$sitzung<-as.numeric(txt2$sitzung)
  txt2$year<-1:dim(txt2)[1]
  
  txt2[txt2$wahlperiode=="17" & txt2$sitzung > 0   & txt2$sitzung < 10  , "year"] <-1
  txt2[txt2$wahlperiode=="17" & txt2$sitzung > 66  & txt2$sitzung < 76 , "year"] <-2
  txt2[txt2$wahlperiode=="17" & txt2$sitzung > 135 & txt2$sitzung < 145 , "year"] <-3
  txt2[txt2$wahlperiode=="17" & txt2$sitzung > 201 & txt2$sitzung < 211 , "year"] <-4
  
  txt2[txt2$wahlperiode=="18" & txt2$sitzung > 0   & txt2$sitzung < 10  , "year"] <-5
  txt2[txt2$wahlperiode=="18" & txt2$sitzung > 60  & txt2$sitzung < 70 , "year"] <-6
  txt2[txt2$wahlperiode=="18" & txt2$sitzung > 130 & txt2$sitzung < 140 , "year"] <-7
  txt2[txt2$wahlperiode=="18" & txt2$sitzung > 196 & txt2$sitzung < 206 , "year"] <-8

  txt2$E_jt<-c(10)
  txt2[txt2$wahlperiode=="17" & txt2$sitzung > 0   & txt2$sitzung < 10  , "E_jt"] <-1
  txt2[txt2$wahlperiode=="17" & txt2$sitzung > 66  & txt2$sitzung < 76 , "E_jt"] <-2
  txt2[txt2$wahlperiode=="17" & txt2$sitzung > 135 & txt2$sitzung < 145 , "E_jt"] <-3
  txt2[txt2$wahlperiode=="17" & txt2$sitzung > 201 & txt2$sitzung < 211 , "E_jt"] <-4
  
  txt2[txt2$wahlperiode=="18" & txt2$sitzung > 0   & txt2$sitzung < 10  , "E_jt"] <-1
  txt2[txt2$wahlperiode=="18" & txt2$sitzung > 60  & txt2$sitzung < 70 , "E_jt"] <-2
  txt2[txt2$wahlperiode=="18" & txt2$sitzung > 130 & txt2$sitzung < 140 , "E_jt"] <-3
  txt2[txt2$wahlperiode=="18" & txt2$sitzung > 196 & txt2$sitzung < 206 , "E_jt"] <-4
  
txt2<-select(txt2, -wahlperiode)
txt2<-select(txt2, -sitzung)
txt2<-txt2[!txt2$E_jt==10,]

print(table(txt2$E_jt))

rm(txt_17)
rm(txt_18)
```


## Select Speaker
In the next step, we reduce the number of speakers: <br>
Only politicians who spoke in all four years and from whom enough speech data (more than 500 sentences) are available, are included.
```{r speaker, echo=FALSE, cache=TRUE}
#speaker that spoke in all four years of a session (E_ij=1:4)
txt2_t <- aggregate(text ~ E_jt + speaker_fp, #group speeches by E_jt = session and speaker, right?
                  data = txt2, 
                  paste, 
                  collapse = " ")

df<-as.data.frame(as.matrix(table(txt2_t$speaker_fp)))
df$names<-rownames(df)
txt2<-txt2[txt2$speaker_fp %in% df$names[df$V1==4],] 
#length(df$names[df$V1==4]) 
  
#select speakers with the highest number of sentences
txt2$ns<-nsentence(txt2$text)
txt2_t2<- aggregate(txt2$ns, by=list(Category=txt2$speaker_fp), FUN=sum)
txt2<-txt2[txt2$speaker_fp  %in% txt2_t2$Category[txt2_t2$x>500],]
#length(txt2_t2$Category[txt2_t2$x>500]) #number of speakers included: 281

#replace ???
txt2$text<-chartr("????", "aous", txt2$text)
txt.c<-corpus(txt2)
```

```{r speaker_head, include=TRUE, cache=TRUE}
print(head(unique(txt2$speaker_fp)))
```
Overall, `r length(unique(txt2$speaker_fp))` speakers are included in the df.

## Prepare 

- Next, we remove punctuation, numbers, capitalization, tokenize the text into sentences and words, and reduce words to their dictionary root. 
- Subsequently, the number of tokens is reduced by removing stop-words, procedural vocabulary and other non-policy words. 
- The list of which words to exclude is adopted from @Ash2017, Appendix, pp.27. 

```{r prepare, include=FALSE, cache=TRUE}
txt.t<-tokens(txt.c,
              remove_punct = TRUE,
              remove_numbers = TRUE,
              what = "word",
              include_docvars= TRUE, 
              remove_symbols = TRUE)
txt.t<-tokens_remove(txt.t, stopwords("german"))
```

```{r prepare_2, include=TRUE, cache=TRUE}
print(sum(ntoken(txt.t)))
source('scr_R/toexclude.R')
txt.t<-tokens_select(txt.t, toexclude, selection = "remove", padding = FALSE)
print(sum(ntoken(txt.t)))
```

```{r prepare_3, include=FALSE, cache=TRUE}
txt.t<-tokens_wordstem(txt.t, language = "de")
txt.t<-tokens_tolower(txt.t)
txt.t<-tokens_remove(txt.t, "?") 
txt.t<-tokens_remove(txt.t, "??")


```

## Bi- and Trigrams
Based on the tokenized speeches, we construct bi- and trigrams. 
To increase computational speed, we chose only phrases with a minimum document and termfrequency of 10.

```{r bitri, include=TRUE, cache=TRUE}
bitri<- tokens_ngrams(txt.t, 1:2) #generate ngrams 
sum(ntoken(bitri))

phi_ikt<-dfm_trim(dfm_group(dfm(bitri), groups = interaction(E_jt, speaker_fp, speaker_party)),
         min_docfreq = 10,
         min_termfreq = 10)
k<-colnames(phi_ikt)
```

## Construct frequency-weighted phrase divisiveness 

$Y_{it}$ measures the divisiveness of speech.

$$
Y_{it} = log(\sum_{k=1}^{K}{\frac{f_{ikt} \cdot \chi_{kt}^2}{F_{it}}})
$$

---

### Political Divisiveness Metric 

- $\chi^2_{kt}$, denotes the political divisiveness of a phrase.
- According to @Gentzkow2010, this metric ranks phrases by their association with particular political parties.

It is constructed as follows. 

- $n_{kt}^{CDU}$: the number of times phrase $k$ is used by CDU/CSU-politicians during session $t$. 
- $N_{t}^{CDU} = \sum_{k}n_{kt}^{CDU}$ are summed frequencies of all phrases used by CDU/CSU-politicians in session $t$. 
- $\tilde{n}_{kt}^{CDU} = N_{t}^{CDU} - n_{kt}^{CDU}$: total number of times phrases besides $k$ were used by CDU/CSU-politicians. 

Pearson’s $\chi_{kt}^2$ statistic for each phrase $k$ at time $t$ is constructed as:

$$
\chi_{kt}^2 = \frac{ ( n_{kt}^{CDU} \tilde{n}_{kt}^{SPD} - n_{kt}^{SPD} \tilde{n}_{kt}^{CDU} )  }
{N_{t}^{CDU} N_{t}^{SPD} ( n_{kt}^{CDU}  + \tilde{n}_{kt}^{SPD} ) ( n_{kt}^{CDU} + \tilde{n}_{kt}^{SPD} ) }
$$

- $\chi_{kt}^2$ provides a test statistic for the null hypothesis that phrase $k$ is used equally by CDU/CSU and SPD-politicians during session $t$
- If the frequencies $n_{kt}^{CDU}$ and $n_{kt}^{SPD}$ are drawn from multinomial distributions.
- $\chi_{kt}^2$ scores language as divisive if it is used more often by just one of the political parties.

```{r chi2_kt, include=FALSE, cache=TRUE, warning=FALSE}
#construct
dfm_kt.f<-dfm(bitri, groups = c("E_jt", "speaker_party")) #group by party
dfm_kt<-dfm_select(dfm_kt.f, k, selection = "keep") #choose same ngrams

#CDU
df.cdu<-as.data.frame(dfm_subset(dfm_kt, substrRight(rownames(dfm_kt), 3)=="csu"))
df.cdu<-df.cdu[,-1] #psi_kt
N_cdu<-rowSums(df.cdu)
```

---

We start constructing $\chi_{kt}^2$ by calculating $N_{cdu}$ and $N_{spd}$

```{r chi2_kt_2, include=TRUE, cache=TRUE, warning=FALSE}
df.spd<-as.data.frame(dfm_subset(dfm_kt, substrRight(rownames(dfm_kt), 3)=="spd"))
df.spd<-df.spd[,-1]
N_spd<-rowSums(df.spd)
print(N_spd)
```

---

$\chi_{kt}^2$ is then constructed as follows:
```{r chi2_kt_prep, include=FALSE, cache=TRUE}
chi2<-as.data.frame(matrix(NA, 
                           nrow = dim(df.cdu)[1], 
                           ncol = dim(dfm_kt)[2]))
rownames(chi2)<-1:4
colnames(chi2)<-colnames(dfm_kt)
```

```{r chi2_kt_loop, include=TRUE, cache=TRUE}
for(i in 1:dim(chi2)[1]){
  for(j in 1:dim(chi2)[2]){
    n_cdu<-df.cdu[i,j]
    n_spd<-df.spd[i,j]

    nt_cdu<-(N_cdu[i]-n_cdu)
    nt_spd<-(N_spd[i]-n_spd)
    
    chi2[i,j]<- (((n_cdu*nt_spd-n_spd*nt_cdu)^2)/(N_cdu[i]*N_spd[i]*(n_cdu+n_spd)*(nt_cdu+nt_spd)))
  }
}
print(chi2[,1:5])
```

---

According to $\chi_{kt}^2$, the 10 most and least divisive phrases in the dataset are:

```{r most_least_phrases, echo=FALSE, cache=TRUE}
chi2[dim(chi2)[1]+1,]<-colSums(chi2)

chi22<-as.data.frame(t(chi2[dim(chi2)[1]+1,]))
chi22$names<-rownames(chi22)
chi22<-chi22[order(-chi22[,1]), ]

chi2<-chi2[-5,]

chi2_t<-cbind(as.matrix(sort(head(chi22$names, 10))),
      as.matrix(sort(tail(chi22$names, 10))))
colnames(chi2_t)<-c("Most Divisive Phrases",
                "Least Divisive Phrases")
chi2_t
```

---

### Normalized Frequency

- Just a reminder: Our goal is to calculate $Y_{it}$

$$
Y_{it} = log(\sum_{k=1}^{K}{\frac{f_{ikt} \cdot \chi_{kt}^2}{F_{it}}})
$$

- We just calculated $\chi_{kt}^2$
- Now, we proceed with $f_{ikt}$
- The normalized frequency $f_{ikt}$ is constructed such that it shows zero mean and standard deviation one (z-score centering, function `scale()`). 
- This means that each phrase has the same influence on the divisiveness measure. 
$$
f_{ikt}:=\frac{\phi_{ikt}-\mu_{kt}}{\sigma_{kt}}
$$

----

$$
f_{ikt}:=\frac{\phi_{ikt}-\mu_{kt}}{\sigma_{kt}}
$$

That's $\phi_{ikt}$:
```{r f_ikt_1, include=TRUE, cache=TRUE, warning=FALSE}
phi_ikt[1:5,1:5]
```

---

$$
f_{ikt}:=\frac{\phi_{ikt}-\mu_{kt}}{\sigma_{kt}}
$$

That's $\mu_{kt}$:
```{r f_ikt_2, echo=FALSE, cache=TRUE, warning=FALSE}
tik<-do.call(rbind.data.frame,
              strsplit(rownames(phi_ikt), split='.', fixed=TRUE))
mu_ikt<-as.data.frame(phi_ikt)
mu_ikt$t<-tik[,1]
mu_ikt$c<-tik[,3]
mu_ikt<-mu_ikt[,-1]
mu_kt<- mu_ikt %>%
  group_by(t,c) %>%
  dplyr::summarise(across(.cols = everything(), 
                   ~ mean(.x, na.rm = TRUE)))
mu_kt[,1:5]
```

---

$$
f_{ikt}:=\frac{\phi_{ikt}-\mu_{kt}}{\sigma_{kt}}
$$

That's $\sigma_{ikt}$
```{r f_ikt_3, include=TRUE, cache=TRUE, warning=FALSE}
sigma_ikt<-as.data.frame(phi_ikt)
sigma_ikt$t<-tik[,1]
sigma_ikt$c<-tik[,3]
sigma_ikt<-sigma_ikt[,-1]
sigma_kt<- sigma_ikt %>%
  group_by(t,c) %>%
  dplyr::summarise(across(.cols = everything(), 
                   ~ sd(.x, na.rm = TRUE)))
sigma_kt[,1:5]
```

---

And here we finally have $f_{ikt}$

$$
f_{ikt}:=\frac{\phi_{ikt}-\mu_{kt}}{\sigma_{kt}}
$$
```{r f_ikt_4, include=TRUE, cache=TRUE, warning=FALSE}
f_ikt <- as.data.frame(t(apply(phi_ikt, 1, scale))) #mean over rows
colnames(f_ikt)<-colnames(phi_ikt)
f_ikt[1:5,1:5]
```

---

### Total Number of Phrases Spoken

- Remember our goal is to calculate $Y_{it}$

$$
Y_{it} = log(\sum_{k=1}^{K}{\frac{f_{ikt} \cdot \chi_{kt}^2}{F_{it}}})
$$

- We started with $Y_{it}$, went on to calculate $f_{ikt}$; now, only $F_{it}$ is left
- $F_{it}$ is the total number of phrases spoken by member $i$ during $t$.
$$
F_it=\sum_{k=1}^{K} f_{ikt}
$$

```{r F_it , echo=FALSE, cache=TRUE}
F_it<-as.data.frame(rowSums(f_ikt))
colnames(F_it)<-"F_it"
head(F_it)
```

## Politician Divisiveness
Politician divisiveness $Y_{it}$ is defined as the frequency-weighted phrase divisiveness for the phrases used by the member.
$$
Y_{it} = log(\sum_{k=1}^{K}{\frac{f_{ikt} \cdot \chi_{kt}^2}{F_{it}}})
$$


```{r Y_it, echo=FALSE, cache=TRUE}
#construct empty df
Y_it<- data.frame(matrix(ncol = 1, nrow = dim(F_it)[1]))
rownames(Y_it)<-rownames(F_it)
colnames(Y_it)<-"y_it"  

#calculate
f_it<-c()
for(i in 1:dim(Y_it)[1]){
  chi<-chi2[substr(rownames(f_ikt)[i], start = 1, stop = 1) ,] # get right E_ij
  for(j in 1:dim(f_ikt)[2]){
    f_it[j]<-((chi[[j]]*f_ikt[i,j])/F_it[i,1])  
  }
  Y_it[i,1]<-log(abs(sum(f_it, na.rm=TRUE))) 
}
```

## Plot
In the followig we plot a histogram of $Y_{it}$
```{r plt_hist, echo=FALSE, cache=TRUE, fig.align='center', warning=FALSE}
Y_it$E_ij<-substr(rownames(Y_it), start = 1, stop = 1)
Y_it$speaker<-substring(rownames(Y_it), 3)

ggplot(Y_it, aes(y_it)) +
geom_histogram()
```

---

- The following figure plots average politicians' speech divisiveness over the course of the four years of a "Legislaturperiode". 
- Error spikes indicate the 95% confidence interval. 
```{r plt_, echo=FALSE, cache=TRUE, warning=FALSE, fig.align='center'}
tgc <- summarySE(Y_it, measurevar="y_it", groupvars="E_ij")
ggplot(tgc, aes(x=E_ij, y=y_it)) + 
    geom_errorbar(aes(ymin=y_it-ci, ymax=y_it+ci), width=.1) +
    geom_line() +
    geom_point()
```

```{r regression, include=FALSE, cache=TRUE}
txt2$ws<-paste(txt2$year, txt2$speaker_fp, txt2$speaker_party, 
               sep=".") 
txt2_r<-txt2[!duplicated(txt2[,"ws"]),] #delete rows where speaker spoke twice in a year todo I should concat them!
txt2_r<-txt2_r[txt2_r$ws %in% rownames(Y_it),]

table(unique(txt2_r$ws))
table(unique(rownames(Y_it)))
sd<-setdiff(rownames(Y_it),txt2_r$ws)
Y_it$names<-rownames(Y_it)

txt2_r[txt2_r$speaker_party=="CDUCSU"] <- 1
txt2_r[txt2_r$speaker_party=="SPD"]    <- 2
```

## Fixed Effects Regression
$$
		Y_{it} = \alpha_{it} + \rho_EE_{it} + \epsilon_{it} 
$$
	
-	$E_{it}$: election treatment
-	$Y_{it}$: divisiveness 
-	$\alpha_{it}$: fixed effects (party-year) 
- $\epsilon_{it}$: omitted variables

---

$$
		Y_{it} = \alpha_{it} + \rho_EE_{it} + \epsilon_{it} 
$$

```{r fe, echo=FALSE, cache=TRUE}

Y_it$year<-txt2$year[match(Y_it$speaker,
                            paste(txt2$speaker_fp, txt2$speaker_party,sep="."))]
Y_it$party<-do.call(rbind.data.frame,
              strsplit(Y_it$speaker, split='.', fixed=TRUE))[,2]

Y_it2<-Y_it %>% tidyr::drop_na()
model.c <- felm(y_it ~ E_ij|factor(party) , data = Y_it2,
                exactDOF=TRUE)
summary(model.c)
```

## Linear Model
```{r linear_model, include=TRUE, cache=TRUE}
reg1<-lm(Y_it$y_it~Y_it$E_ij)
summary(reg1)
```
## References
